On modeling for Kerr black holes: Basis learning, QNM frequencies, and spherical-spheroidal mixing coefficients

### 2018-07-27 docs/mvfits_paper/mvf.tex
updating paper

\par In particular, a perturbed Kerr \bh{} (e.g. resulting from \bbh{} merger) will have \gw{} radiation that rings down with characteristic frequencies,

@@

When considering many noise realizations, we find that sample noise and residuals have a typically correlation of $99.46\%$.

@@

\end{table}
%
%
\par In this example case, \gmvr{} correctly recovers the functional form of the input data. But, in general, \gmvr{} and related techniques will attempt to model minor correlations and offsets within the training data's noise.
%
However, we have demonstrated the utility of \gmvr{} in a relatively ideal usage case where the underlying function is rational, and the training data is inly weakly contaminated with noise.
%
In the following sections, we consider cases where the functional form of the sample data is not known to be explicitly rational, but the amount of noise within the training data is negligible.
%
%
\begin{figure*}
  %
  \includegraphics[width=\textwidth]{fig/issue3_summary.pdf}
  %
	\caption{ Standard summary plot for \gmvr. }
  %
  \label{fig:gmvrtoy}
  %
\end{figure*}

%
\begin{figure}[htb]
  %
  \includegraphics[width=0.45\textwidth]{fig/issue3_pgreedy.pdf}
	\caption{ Convergence of greedy process for \gmvr{} toy problem. }
  %
\end{figure}

%
\subsection{Modeling \qnm{} frequencies using \gmvp{}}
%
In seeking to apply \gmvp{} to select \qnm{} frequencies, we wish to impose a zero-damping constraint, namely that some frequencies are real as $\jf \rightarrow 1$.
%
We also wish the impose a domain transformation, $\kappa(\jf)$, such that $0 \leq \kappa \leq 1$ and the individual \qnm{} frequencies are morphologically simplified.
%
\par For the domain transformation, inspection of \qnm{} frequencies with $\ell \leq 5$ suggest that
%
\begin{align}
  \kappa(\jf) = \left( \ln( 2 - \jf ) / \ln(3) \right)^{1/(2-\ell+|m|)}
\end{align}
%
appropriately linearizes the sharp behavior of each frequency near $\jf=1$.
%
\par For the zero-damping constraint, when considering a \qnm{} frequency $\cw_{\lmn}$, zero-damping at $\jf=1$ implies that $\cw_{\lmn}(\kappa \approx 0)-m/2 \propto \kappa$, where $m/2$ is the well known limiting value for each \qnm{} frequencies real part as $\jf\rightarrow 1$.
%
% Plots of QNM frequencies
\begin{figure*}
  %
  \begin{tabular}{lcr}
    \includegraphics[width=\figfactor\textwidth]{fig/fits_w.pdf} & \includegraphics[width=\figfactor\textwidth]{fig/fits_tau.pdf}
  \end{tabular}
  %
	\caption{ Fits of dimensionless \qnm{} central frequencies (solid lines) along with select numerical values (grey markers) computed using Leaver's method \cite{Leaver85}.
  %
  Before the application of $\kappa(\j{})$, points are spaced between -\CwFitCalibrationRegion and \CwFitCalibrationRegion according to \CwFitCalibrationRegion times the $\sin$ of a fiducial angle which is uniformly spaced between $-\pi/2$ and $\pi/2$. Values of $\j{}$ are shown in the upper axis for $\kappa$ at $l=m$.
  %
  The grey dashed line marks the value of $\kappa$ where $\j{}=0$. Fits of dimensionless \qnm{} decay rates (solid lines) along with select numerical values (grey markers) computed using Leaver's method \cite{Leaver85}. }
  %
  \label{fig:qnms}
  %
\end{figure*}
%
% Plots of QNM frequencies
\begin{figure*}
  %
  \begin{tabular}{lcr}
    \includegraphics[width=\figfactor\textwidth]{fig/res_w.pdf} & \includegraphics[width=\figfactor\textwidth]{fig/res_tau.pdf}
  \end{tabular}
  %
	\caption{ Fractional residual errors for fits of dimensionless \qnm{} central frequencies and decay times (solid lines) along with select numerical values (grey markers) computed using Leaver's method \cite{Leaver85}. }
  %
  \label{fig:qnm_err}
  %
\end{figure*}
%
This implies that
%
\begin{align}
  \label{eq:zd}
  \cw_{\lmn} = m/2 + \kappa \sum_{j=1}^{J} c_j \kappa^k \; .
\end{align}
%
In the case of the non-zero damped \qnm{s}, (e.g. $(\ell,m)=(3,2)$), a more general polynomial form may be adopted, namely,
%
\begin{align}
  \label{eq:nzd}
  \cw_{\lmn} = \sum_{j=0}^{J} c_j \kappa^k \; .
\end{align}
%
\par The polynomial content of \eqn{eq:zd} and \eqn{eq:nzd} is determined by \gmvp{}.
%
\eqns{eq:cw_fit_1}{eq:cw_fit_9} display the resulting polynomial models.
%
Although each model's fractional error is within 2\%, each is error is dominated by well known nonlinear oscillations which are considered to be outside of the experimental accuracy of current \gw{} efforts.
%
\par \fig{fig:qnms} displays select training points, as well as model fits for $\cw_{lmn}$'s real and imaginary parts.
%
For the right and left panels, simple polynomial behavior of each curve is a result of the displayed linear domain in $\kappa(\jf)$.
%
In the left panel, we have scaled $1/\tau_{\lmn}$ by factors for $m/(n+1/2)$ to place the \qnm{s} with $n=0$ and $n=1$ at approximately the same scale.
%
% Equations for QNM frequency fits
\begin{widetext}
	\input{src/cwfit_eqns}
\end{widetext}
%
%
\subsection{Modeling spherical-spheroidal inner-products using \gmvr{}}
%
%
% Equations for mixing coefficients
\begin{widetext}
  \input{src/ysprod_eqns}
\end{widetext}

>>>>

\par In particular, a perturbed Kerr \bh{} (e.g. resulting from \bbh{} merger) will have \gw{} radiation that rings down with characteristic \textit{dimensionless} frequencies,

@@

When considering many noise realizations, we find that sample noise and residuals have an average correlation of $99.46\%$.

@@

\end{table}
%
%
\par In this example case, \gmvr{} correctly recovers the functional form of the input data. But, in general, \gmvr{} and related techniques, having no knowledge of , will attempt to model minor correlations and offsets within the training data's noise.
%
However, we have demonstrated the utility of \gmvr{} in a relatively ideal usage case where the underlying function is rational, and the training data is inly weakly contaminated with noise.
%
In the following sections, we consider cases where the functional form of the sample data is not known to be explicitly rational, but the amount of noise within the training data is negligible.
%
%
\begin{figure*}
  %
  \includegraphics[width=\textwidth]{fig/issue3_summary.pdf}
  %
	\caption{ Standard summary plot for \gmvr. }
  %
  \label{fig:gmvrtoy}
  %
\end{figure*}
\begin{figure}[htb]
  \includegraphics[width=0.45\textwidth]{fig/issue3_pgreedy.pdf}
	\caption{ Convergence of greedy process for \gmvr{} toy problem. }
\end{figure}
% %%%%%%
\subsection{Modeling \qnm{} frequencies using \gmvp{}}
% %%%%%%
In seeking to apply \gmvp{} to select \qnm{} frequencies, we wish to account for the known extremal Kerr behavior of some modes.
Namely, we will impose a zero-damping constraint: some frequencies are real as $\jf \rightarrow 1$.
We also wish the impose a domain transformation, $\kappa(\jf,\ell,m)$, such that $0 \leq \kappa \leq 1$ and the individual \qnm{} frequencies are made approx. polynomial in $\kappa$.
\par For the domain transformation, inspection of \qnm{} frequencies with $\ell \leq 5$ suggest that
  \kappa(\jf,\ell,m) = \left( \log_3( 2 - \jf ) \right)^{1/(2-\ell+|m|)}
appropriately linearizes the sharp behavior of each frequency near $\jf=1$, and maps $-1 \le \jf \le 1$ onto $0 \le \kappa \le 1$.
\par Towards the zero-damping constraint, when considering a \qnm{} frequency $\cw_{\lmn}$, zero-damping at $\jf=1$ implies that $\cw_{\lmn}(\kappa \approx 0)-m/2 \propto \kappa$, where $m/2$ is the well known limiting value for each \qnm{} frequencies real part as $\jf\rightarrow 1$.
% Plots of QNM frequencies
\begin{figure*}
  \begin{tabular}{lcr}
    \includegraphics[width=\figfactor\textwidth]{fig/fits_w.pdf} & \includegraphics[width=\figfactor\textwidth]{fig/fits_tau.pdf}
	\caption{ Fits of dimensionless \qnm{} central frequencies (solid lines) along with select numerical values (grey markers) computed using Leaver's method \cite{Leaver85}.
  Before the application of $\kappa(\j{})$, points are spaced between -\CwFitCalibrationRegion and \CwFitCalibrationRegion according to \CwFitCalibrationRegion times the $\sin$ of a fiducial angle which is uniformly spaced between $-\pi/2$ and $\pi/2$. Values of $\j{}$ are shown in the upper axis for $\kappa$ at $l=m$.
  The grey dashed line marks the value of $\kappa$ where $\j{}=0$. Fits of dimensionless \qnm{} decay rates (solid lines) along with select numerical values (grey markers) computed using Leaver's method \cite{Leaver85}. }
  \label{fig:qnms}
\end{figure*}
% Plots of QNM frequencies
\begin{figure*}
  \begin{tabular}{lcr}
    \includegraphics[width=\figfactor\textwidth]{fig/res_w.pdf} & \includegraphics[width=\figfactor\textwidth]{fig/res_tau.pdf}
	\caption{ Fractional residual errors for fits of dimensionless \qnm{} central frequencies and decay times (solid lines) along with select numerical values (grey markers) computed using Leaver's method \cite{Leaver85}. }
  \label{fig:qnm_err}
\end{figure*}
  \label{eq:zd}
  \cw_{\lmn} = m/2 + \kappa \sum_{j=1}^{J} c_j \kappa^k \; .
In the case of the non-zero damped \qnm{s}, (e.g. $(\ell,m)=(3,2)$), a more general polynomial form may be adopted, namely,
  \label{eq:nzd}
  \cw_{\lmn} = \sum_{j=0}^{J} c_j \kappa^k \; .
\par The polynomial content of \eqn{eq:zd} and \eqn{eq:nzd} is determined by \gmvp{}.
\eqns{eq:cw_fit_1}{eq:cw_fit_9} display the resulting polynomial models.
In particular, the domain map allows most \qnm frequencies to be well modeled by 4th order polynomials which include all lower degree terms; concurrently, the real and imaginary parts of each $\cw\lmn$ are modeled simultaneously.
Although each model's fractional error is within 2\%, each is error is dominated by well known nonlinear oscillations which are considered to be outside of the experimental accuracy of current 2nd generation \gw{} efforts.
\par \fig{fig:qnms} displays select training points, as well as model fits for $\cw\lmn$'s real and imaginary parts.
For the right and left panels, simple polynomial behavior of each curve is a result of the displayed linear domain in $\kappa(\jf,\ell,m)$.
In the left panel, we have scaled $1/\tau\lmn$ by factors for $m/(n+1/2)$ to place the \qnm{s} with $n=0$ and $n=1$ at approximately the same scale.
% Equations for QNM frequency fits
\begin{widetext}
	\input{src/cwfit_eqns}
\end{widetext}
\par In each panel of \fig{fig:qnms}, the horizontal upper and lower demonstrate the effect of mapping $\jf$ onto $\kappa$.
In particular, it is shown that the two solution branches ($\jf>0$ and $\jf<0$) naturally form a single family of solutions when accounting for the sign of the \bh{}'s oriented spin.
Concurrently, the use of $\kappa$ as a domain variable has the desirable effect of making each $\cw\lmn$ and $\tau\lmn$ approximately parabolic.
We note that, in the asymptotic vicinity of $\jf=1$, the \qnm{} frequencies and decay times are known to have solutions that are asymptoticly degenerate.
In allowing \eqns{eq:cw_fit_1}{eq:cw_fit_9} to extrapolate to $\jf \sim 1$, we do not explicitly account for this additional effect.
\subsection{Modeling spherical-spheroidal inner-products using \gmvr{}}
% Equations for mixing coefficients
\begin{widetext}
  \input{src/ysprod_eqns}
\end{widetext}
